Convective heat transport in stratified atmospheres at low and high Mach number

### 2016-10-13 tex/nusselt_paper.tex
shapes up the discussion a bit

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/ma_v_eps.png}
\caption{(a) Characteristic horizontally averaged maximum Mach numbers have been
time averaged for over $\geq 100t_{b}$ beginning roughly 50$t_{b}$ after the start of simulations. For $\epsilon \leq 0.1$,
an evident scaling of $Ma \propto \{\epsilon^{0.50}, \epsilon^{0.51}\}$ 
at Ra = $\{10^2, 10^5\}$ is retrieved.
When $\epsilon \rightarrow m_{ad}$, large deviations from this power law are seen and the systems display
supersonic dynamics.  
(b) At high epsilon, Ma scales as Ra$^{0.26}$ until it reaches the supersonic regime, at which point it turns
over and follows a power law of Ra$^{-0.10}$.  At low epsilon, consistent power laws are achieved throughout all
values of $Ra$ studied, where Ma $\propto$ \{Ra$^{0.26}$, Ra$^{0.22}$\} for $\epsilon = \{10^{-4}, 10^{-7}\}$. All
error bars are negligible.  \label{fig:ma_v_eps} }
\end{figure}

@@

After appropriately time-averaging the fluxes, a sensible Nusselt number is retrieved.  Nusselt numbers for
all simulations at low and high Ma are plotted in Fig. \ref{fig:nu_v_ra}.  At $\epsilon = \{10^{-4}, 10^{-7}\}$,
scaling laws of $Nu \propto Nu^{\{0.31, 0.31\}}$ are retrieved.  At $\epsilon = 0.5$, in the near-sonic
regime ($Ra \leq 10^4$), the scaling of $Nu$ with $Ra$ is inflated, with $Nu \propto Ra^{0.45}$.  As simulations
pass into the supersonic regime and shocks start to form near the downflows, propoelling warm fluid deep
into the atmosphere, that scaling drops to $Nu \propto Ra^{0.19}$.  Error on all power law scalings is
negligible.

@@

\section{Discussion}
\refstepcounter{section}
\label{sec:discussion}
In this letter we have studied fundamental heat transport by stratified convection in simplified 2-D polytropic
atmospheres at low and high Mach number.  Out atmospheres here are specified by two parameters, \nrho 
and $\epsilon$.  We argue that these atmospheres are the natural extension
of the RB problem to stratified systems, and should be used to understand the basic properties of stratified
convection.  The striking similarity between the growth of Nu with Ra at low Ma and in RB convection suggests
that low Ma polytropic convection shares many properties with RB convection but includes the complicating
effect of stratification. We have argued that the traditional definition of the Nusselt number in stratified atmospheres
\cite{graham1975} is the correct one, and shown how it extends in extreme cases.

Furthermore, we have demonstrated that low Ma flows exhibit the same broad-upflow / narrow downflow dynamics
as are observed in high Mach number flows \cite{hurlburt&all1984}.  At low Ma, flows are in pressure equilibrium
with their surroundings, so cold parcels must also be low density parcels, both of which are negative
entropy contributions.  As there is no thermodynamic variable causing additional breaking or acceleration, it
is simply the effects of stratification which lead downflows to be narrow and upflows to be broad.
It is possible that at low Ma, these flows begin to feel the pressure response of the
top and bottom walls which leads them to deflect at the far boundary in a way that high Ma flows do not.

The dynamics of these polytropic solutions are complex and highly time-dependent, even in two dimensions.
Time-dependent oscillating shear states have developed spontaneously, as seen before in RB convection
\cite{goluskin&all2014}.  Future work will aim to better understand the mechanisms of shearing states and
whether or not these states are attainable in three-dimensional, non-rotating atmospheres.  While the
two-dimensional work studied here offers a basis for comparing the heat transport abilities of stratified
convection compared to RB convection \cite{johnston&doering2009}, it will be informative to study the low 
Mach number systems presented here in three dimensions and in atmospheres bounded by stable regions
\cite{hurlburt&all1986}.

The similarity between the scaling of Nu in RB convection and in these stratified polytropes suggests that
a similar Prandtl-Blausius or Grossmann-Lohse boundary layer theory could be developed for fully compressible
convection in these systems \cite{ahlers&all2009}.  If our scalings of Ma and $\epsilon$ hold up, then under
solar conditions (Ra $\approx 10^{20}$, Ma $\approx 10^{-4}$), we expect that $\epsilon \approx 10^{20}$ and
Nu $\approx 10^{6}$.  Solar conditions are of course more complicated, as there $\kappa$ is set by the
radiative opacity and depend on both $T$ and $\rho$.  Our studies here set the groundwork for studying transport
under conditions with more realistic $\kappa$.

>>>>

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/ma_v_eps.png}
\caption{(a) Characteristic horizontally averaged maximum Mach numbers have been
time averaged for over $\geq 100t_{b}$ beginning roughly 50$t_{b}$ after the start of simulations. For $\epsilon \leq 0.1$,
an evident scaling of $Ma \propto \{\epsilon^{0.50}, \epsilon^{0.51}\}$ 
at Ra = $\{10^2, 10^5\}$ is retrieved.
When $\epsilon \rightarrow m_{ad}$, large deviations from this power law are seen and the systems display
supersonic dynamics.  
(b) At high epsilon, Ma scales as Ra$^{0.28}$ until it reaches the supersonic regime, at which point it turns
over and follows a power law of Ra$^{-0.10}$.  At low epsilon, consistent power laws are achieved throughout all
values of $Ra$ studied, where Ma $\propto$ \{Ra$^{0.26}$, Ra$^{0.22}$\} for $\epsilon = \{10^{-4}, 10^{-7}\}$. All
error bars are negligible.  \label{fig:ma_v_eps} }
\end{figure}

@@

After appropriately time-averaging the fluxes, a sensible Nusselt number is retrieved.  Nusselt numbers for
all simulations at low and high Ma are plotted in Fig. \ref{fig:nu_v_ra}.  At $\epsilon = \{10^{-4}, 10^{-7}\}$,
scaling laws of $Nu \propto Nu^{\{0.31, 0.31\}}$ are retrieved.  At $\epsilon = 0.5$, in the near-sonic
regime ($Ra \leq 10^4$), the scaling of $Nu$ with $Ra$ is inflated, with $Nu \propto Ra^{0.45}$.  As simulations
pass into the supersonic regime and shocks start to form near the downflows, propelling warm fluid deep
into the atmosphere, that scaling drops to $Nu \propto Ra^{0.19}$.  Error on all power law scalings is
negligible.

@@

\section{Discussion}
\refstepcounter{section}
\label{sec:discussion}
In this letter we have studied fundamental heat transport by stratified convection in simplified 2-D polytropic
atmospheres at low and high Mach number.  Our atmospheres here are specified by two parameters, \nrho 
and $\epsilon$.  We argue that these atmospheres are the natural extension
of the RB problem to stratified systems, and should be used to understand the basic properties of stratified
convection.  The similarity between the scaling of Nu in RB convection and in these stratified polytropes suggests that
a boundary layer theory such as the Prandtl-Blausius or Grossmann-Lohse theories for incompressible flows
could be developed for fully compressible
convection in these systems \cite{ahlers&all2009}.

Furthermore, we have demonstrated that low Ma flows exhibit the same pattern of broad upflows and narrow downflows
observed in high Ma flows \cite{hurlburt&all1984}.  At low Ma, flows are often in pressure equilibrium
with their surroundings, so cooling parcels increases their density, and both of these effects reduce
entropy.  Neither of the two free thermodynamic variables can cause breaking in warm upflow regions, 
so stratification effects must cause the observed flow asymmetry.
While not studied in this work, it is possible that at low Ma, flows feel the pressure response of the
boundaries, causing deflection in a manner unseen at high Ma.

The dynamics of these polytropic solutions are complex and highly time-dependent, even in two dimensions.
Time-dependent oscillating shear states have developed spontaneously, as seen before in RB convection
\cite{goluskin&all2014}.  While computationally difficult, the highest values of Ra and the lowest value
of $\epsilon$ studied here are far below the values found in nature.  If the scalings of Nu and Ma
presented here hold, then under solar conditions ($\text{Ra }\approx 10^{20}$, $\text{Ma }\approx 10^{-4}$), we expect that $\epsilon \approx 10^{20}$ and
$\text{Nu }\approx 10^{6}$.

Future work will aim to better understand the mechanisms of shearing states and
whether or not these states are attainable in three-dimensional, non-rotating atmospheres.  Our studies
here have set the groundwork for understanding and comparing heat transport in stratified convection
to that in RB convection \cite{johnston&doering2009}, and for future studies of transport in stratified
convection in more realistic systems, such as those bounded by stable regions \cite{hurlburt&all2009} or 
using more realistic profiles of $\kappa$.
